Fast and scalable Gaussian process modeling of stellar variability \vspace{-3\baselineskip}

### 2016-11-29 paper/ms.tex
building some docs
making a change to the paper

@@

\section{Outline}

\begin{enumerate}
%\item Non-parametric modeling of time series with correlated noise using Gaussian processes
%\item Assumptions:\\
%  a). Stationary noise (although this can be broken).\\
%  b). Gaussian
%\item Problem: can't handle large datasets - e.g. entire Kepler long-cadence; Spitzer light
%   curves; Solar curves; short-cadence.
%\item Solution(s):\\
%  a). HODLR \citep{Ambikasaran2013,Ambikasaran2016}\\
%  b). CARMA (Lorentzian power spectra, but with limitations) \citep{Kelly2014}\\
%  c). Tri-diagonal (Press \& Rybicki) \citep{1995PhRvL..74.1060R}
%  d). KISS-GP \citep{WilsonNickisch2015} %Kernel Interpolation for Scalable Structured Gaussian Processes (KISS-GP) Andrew Wilson, Hannes Nickisch Proceedings of The 32nd International Conference on Machine Learning, pp. 1775-1784, 2015
%\item Problem: HODLR is tricky to use; CARMA requires certain relations between
%  coefficients of Lorentzians (right?); P\&R only works for up to 2 components + white noise
%\item Solution: Ambikasaran \citep{Ambikasaran2015} shows how sum of exponential kernels can be solved in order
%   $O(N*p^2)$, where p is number of Lorentzians.
%  - Sum of Lorentzians is a good description of stellar variability:
%     - granulation has expoential (or sums of exponential) noise
%     - asteroseismic obeys sums of Lorentzians
%     - periodic variability due to spots/plages/etc. can also be
%       approximated by Lorentzian(s)
%\item Generalized Ambikasarn for complex coefficients (Hermitian).
\item Performance:\\
  a). Show $O(N*p^2)$ scaling\\
  b). Show comparison to HODLR.\\
  c). Show comparison to wavelet (ala Carter \& Winn).\\
  d). Compare with CARMA Kallman filter(?).
\item Various methods/formulae:\\
  a). Likelihood computation.\\
  b). Generating GP with particular covariance [yet to solve]. [ ]\\
  c). Inferring particular components.\\
  d). Computing derivatives of GPs (as used in RV methods).\\
  e). Forecasting/interpolating.\\
  f). Inferring confidence intervals.\\
  g). Derivative of likelihood function with respect to kernel parameters.\\
  h). Non-stationary GP (can the coefficients vary with time?).\\
  i). Multi-band (each element of extended matrix becomes a matrix).
  j). Approximating various kernels: Matern, exp(sin$^2$), Gaussian, etc.
\item Generalization to non-stationary noise.
\item Example application: TYC 3559\\
  a). Entire long-cadence 17-quarter Kepler dataset (~65k data points).\\
  b). Analysis of solar data.
  c). CO2 Mauna Kea
  d). TBD Computing derivatives of GPs (as used in RV methods).\\
  e). TBD Forecasting/interpolating.\\
  f). TBD Inferring confidence intervals.\\
  g). Autodiff: Derivative of likelihood function with respect to kernel parameters.\\
  h). TBD Non-stationary GP (can the coefficients vary with time?).\\
  i). TBD Multi-band (each element of extended matrix becomes a matrix).

\item Example application: TYC 3559\\
  a). Entire long-cadence 17-quarter Kepler dataset (~65k data points).\\
  b). Simulated datasets - hyperparameter/PSD inference.\\
  c). Analysis of solar data?

\item Possible future applications:\\
  a). Inference of asteroseismic variability - ala Andrew Gordon Wilson \citep{2013arXiv1302.4245W}, but
      with Lorentzians, not Gaussians.\\
  b). Multi-waveband GPs.\\
  c). RV interpolation.\\
  d). Non-parametric phase functions.\\
  e). Better calibration of the flicker log(g) relation.\\
  f). Measurement of rotational periods - gyrochronology.\\
  g). Doppler shifts in oscillating EBs.

\item Appendix:
  a). Description/summary of mathematics.\\
  b). Description of code.
\end{enumerate}

>>>>

%
% New Outline:
% 1. Introduction
% 2. Gaussian processes
%   - GPs in general
%   - RP method
%   - Ambikasaran method
%   - Generalization
% 3. Implementation and performance
%   - Implementation considerations: band solver vs. sparse, parameterization
%   - Benchmark numbers
%   - Scaling with N and M
%   - Description of APIs
% 4. Examples
%   - Simulated data: a simple exact GenRP simulation, a GenRP simulation with
%     structure, a few standard kernels, a AGW kernel
%   - An AGN light curve
%   - An RR Lyrae
%   - A kepler light curve
%   - An asteroseismic target
% 5. Comparisons to other methods
% 6. Summary
%   - Possible applications
%

\section{Gaussian processes in one-dimension}

\citep{Rasmussen:2006}

\subsection{Gaussian processes with Lorentzian power spectra}

\citep{Rybicki:1995, Ambikasaran:2015}

\subsection{Generalization to quasi-periodic covariance}

Our method.

\section{Implementation \& performance}

\subsection{Implementation considerations}

Solvers.
Banded vs.\ sparse.

\subsection{Benchmarks \& scaling}

\subsection{Parameterization \& API}

\section{Examples}

\subsection{Simulated data}

\begin{enumerate}
\item an OU process
\item a full GenRP process
\item Some standard kernels: Matern, Exp-squared, etc.
\item A KISS-GP process
\end{enumerate}

\subsection{Real data}

\begin{enumerate}
\item AGN
\item RR Lyrae
\item Kepler with simulated transit
\item Asteroseismic target
\end{enumerate}

\section{Comparisons to other methods}

Toeplitz, KISS-GP, CARMA, HODLR.

Limitations of GenRP: one-dimension, stationary, etc.

@@

% \section{Outline}
% \begin{enumerate}
%\item Non-parametric modeling of time series with correlated noise using Gaussian processes
%\item Assumptions:\\
%  a). Stationary noise (although this can be broken).\\
%  b). Gaussian
%\item Problem: can't handle large datasets - e.g. entire Kepler long-cadence; Spitzer light
%   curves; Solar curves; short-cadence.
%\item Solution(s):\\
%  a). HODLR \citep{Ambikasaran2013,Ambikasaran2016}\\
%  b). CARMA (Lorentzian power spectra, but with limitations) \citep{Kelly2014}\\
%  c). Tri-diagonal (Press \& Rybicki) \citep{1995PhRvL..74.1060R}
%  d). KISS-GP \citep{WilsonNickisch2015} %Kernel Interpolation for Scalable Structured Gaussian Processes (KISS-GP) Andrew Wilson, Hannes Nickisch Proceedings of The 32nd International Conference on Machine Learning, pp. 1775-1784, 2015
%\item Problem: HODLR is tricky to use; CARMA requires certain relations between
%  coefficients of Lorentzians (right?); P\&R only works for up to 2 components + white noise
%\item Solution: Ambikasaran \citep{Ambikasaran2015} shows how sum of exponential kernels can be solved in order
%   $O(N*p^2)$, where p is number of Lorentzians.
%  - Sum of Lorentzians is a good description of stellar variability:
%     - granulation has expoential (or sums of exponential) noise
%     - asteroseismic obeys sums of Lorentzians
%     - periodic variability due to spots/plages/etc. can also be
%       approximated by Lorentzian(s)
%\item Generalized Ambikasarn for complex coefficients (Hermitian).
% \item Performance:\\
%   a). Show $O(N*p^2)$ scaling\\
%   b). Show comparison to HODLR.\\
%   c). Show comparison to wavelet (ala Carter \& Winn).\\
%   d). Compare with CARMA Kallman filter(?).
% \item Various methods/formulae:\\
%   a). Likelihood computation.\\
%   b). Generating GP with particular covariance [yet to solve]. [ ]\\
%   c). Inferring particular components.\\
%   d). Computing derivatives of GPs (as used in RV methods).\\
%   e). Forecasting/interpolating.\\
%   f). Inferring confidence intervals.\\
%   g). Derivative of likelihood function with respect to kernel parameters.\\
%   h). Non-stationary GP (can the coefficients vary with time?).\\
%   i). Multi-band (each element of extended matrix becomes a matrix).
%   j). Approximating various kernels: Matern, exp(sin$^2$), Gaussian, etc.
% \item Generalization to non-stationary noise.
% \item Example application: TYC 3559\\
%   a). Entire long-cadence 17-quarter Kepler dataset (~65k data points).\\
%   b). Analysis of solar data.
%   c). CO2 Mauna Kea
%   d). TBD Computing derivatives of GPs (as used in RV methods).\\
%   e). TBD Forecasting/interpolating.\\
%   f). TBD Inferring confidence intervals.\\
%   g). Autodiff: Derivative of likelihood function with respect to kernel parameters.\\
%   h). TBD Non-stationary GP (can the coefficients vary with time?).\\
%   i). TBD Multi-band (each element of extended matrix becomes a matrix).

% \item Example application: TYC 3559\\
%   a). Entire long-cadence 17-quarter Kepler dataset (~65k data points).\\
%   b). Simulated datasets - hyperparameter/PSD inference.\\
%   c). Analysis of solar data?

% \item Possible future applications:\\
%   a). Inference of asteroseismic variability - ala Andrew Gordon Wilson \citep{2013arXiv1302.4245W}, but
%       with Lorentzians, not Gaussians.\\
%   b). Multi-waveband GPs.\\
%   c). RV interpolation.\\
%   d). Non-parametric phase functions.\\
%   e). Better calibration of the flicker log(g) relation.\\
%   f). Measurement of rotational periods - gyrochronology.\\
%   g). Doppler shifts in oscillating EBs.

% \item Appendix:
%   a). Description/summary of mathematics.\\
%   b). Description of code.
% \end{enumerate}
